---
title: "lab_02"
author: "derek willis"
date: "2023-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most?
**A1.** 26 different departments have offered a course with "Climate" in the title that have at least one seat. The department that has offered the most is Atmospheric and Oceanic Science wiht 68.

```{r}
umd_courses |> 
  filter(str_detect(title, "Climate") & seats > 0) |>
  group_by(department) |>
  summarise(classes = n()) |>
  arrange(desc(classes))
```

**Q2.** Of the departments listed in A1, which one is the most interesting to you, and why?
**A2.** I think the most interesting department I found in the answer to A1 was probably economics. A lot of the departments made sense and I think economics does too, but I wonder if the classes offered in this department are interdisciplinary courses relating to the economy and climate change (which sounds interesting) or if the course titles might be using climate in a different sense than relating to weather.


**Q3** Which departments have the most courses in the Fall 2023 term? And which departments in that term have at least one syllabus posted? How are the top five departments different in each result?

**A3.** The Business and Management department has the most courses in the Fall 2023 term. 125 different departments have at least one course listed with a syllabus. The joint-fourth place departments have half the amount of courses with a syllabus compared to the top department, Business and Management.

```{r}
umd_courses |> 
  filter(term == 202308 & syllabus_count > 0) |>
  group_by(department) |>
  summarise(classes = n()) |>
  arrange(desc(classes)) 
  
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as the same dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.
**A4.** The month with by far the most expenses was October, while July and June both featured a lot of spending. In general, spending increased over tim with all months in 2022 placing above months in 2021, but not linearly.

```{r}
maryland_expenses |>
  mutate(expenditure_date = mdy(expenditure_date)) |>
  mutate(month = floor_date(expenditure_date, "month")) |>
  filter(committee_name == 'Moore  Wes For Maryland') |>
  group_by(month) |>
  summarise(total_amount = sum(amount)) |>
  arrange(desc(total_amount)) 
```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for the others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe).
**A5** There was about $26.2 million more dollars spent in Maryland than in other states in the U.S.

```{r}
maryland_expenses |>
  mutate(spending_location = case_when(
        str_detect(address, " Maryland ") ~ "In-state",
        .default = "Out of State"
    )
  ) |>
  group_by(spending_location) |>
  summarise(total_amount = sum(amount,na.rm=TRUE))
```


**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.
**A6.** My pre-lab question was "Which UMD course has the lowest number of seats, not counting classes that have 0 seats associated with them?" Because I saw there were a lot with 1, I wanted to see which department had the most classes with 1 seat. Funnily enough, journalism had by far the most classes with just one seat in them with 52.

```{r}
umd_courses |>
  filter(seats == 1) |>
  group_by(department) |>
  summarise(classes = n()) |>
  arrange(desc(classes)) 
  
```
